---
title: "First steps with ggplot2 library"
output:
  html_notebook: default
  html_document:
    theme: flatly
  pdf_document: default
---

## Types of basic charts and variant on `ggplot`

### Scatterplots

We load the `cars` dataset and plot it.
```{r}
data(cars)
help(cars)
head(cars)
plot(cars)
```


Load the `ggplot` library for observing how it improves the R's basic charts:
```{r}

# install.packages("ggplot2") # Install if it doesn't already be installed 
library(ggplot2)

ggplot(cars, aes(x=speed, y=dist)) + # It loads the data and set the variables
  geom_point(shape=1) # It sets the points like hollow circles


ggplot(cars, aes(x=speed, y=dist)) + 
  geom_point(shape=1) +
  geom_smooth(method=lm) # It adds a regression line with a confidence interval of 95%


ggplot(cars, aes(x=speed, y=dist)) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE) # It adds a regression line without a confidence interval


ggplot(cars, aes(x=speed, y=dist)) + 
  geom_point(shape=1) +
  geom_smooth() # It adds a smooth line above data and a confidence interval of 95%

```



More scatterplots

```{r}
# Load other data
data(diamonds)
help(diamonds)
head(diamonds)
str(diamonds)

```

```{r}
ggplot(diamonds, aes(x=carat, y=price)) + 
  geom_point() # If shape isn't specified, it adds filled circles

ggplot(diamonds, aes(x=carat, y=price)) +
  geom_point(shape=25, size=4) # We can specify the shape and size of points

ggplot(diamonds, aes(x=carat, y=price, colour=cut)) + # Also, we can plot by a factor
  geom_point(size=1.5)

# Even, it can be plot with a gray scale by a value of continue atribute
grayScale <- ggplot(diamonds, aes(x=carat, y=price))
grayScale + geom_point(alpha=1)
grayScale + geom_point(alpha=0.8)
grayScale + geom_point(alpha=0.5)
grayScale + geom_point(alpha=0.3)


# Also, a scatterplot with bins can be do in coloured rectangles
bin <- ggplot(diamonds, aes(x = carat, y = price))
bin + stat_bin2d()
```





### Histograms

We load `mpg` dataset. We plot a histogram with the highway consumption.
```{r}
data(mpg)
help(mpg)
head(mpg)
hist(mpg$hwy)
```


`ggplot` histograms with options of density curve and setting of bin's number.

```{r}
ggplot(mpg, aes(x=hwy)) + 
  geom_histogram(binwidth=3) # It sets the binwidth of each partition

ggplot(mpg, aes(x=hwy)) +
  geom_histogram(bins=30) # It sets the number of partitions

ggplot(mpg, aes(x=hwy)) +
  geom_histogram(binwidth=0.5, colour="red", fill="green") # Colour's options

ggplot(mpg, aes(x=hwy)) +
  geom_density(fill="yellow") # Plot a density curve

# Histogram with a density curve
ggplot(mpg, aes(x=hwy)) +
  geom_histogram(aes(y=..density..), 
                 binwidth=1.2,
                 colour="black", fill="blue") +
  geom_density(alpha=0.3, fill="red")
  
```


Also, histograms can be represented with the variable divided respect of several categories. For it, we go to select two manufacturers: audi and volkswagen:

```{r}
mpg_subset <- mpg[mpg$manufacturer == "audi" | mpg$manufacturer == "volkswagen", ]
str(mpg_subset)
```


```{r}
# Overlapped histograms
ggplot(mpg_subset, aes(x=hwy, fill=manufacturer)) + # With fill=manufacturer, we say that the histogram is filled with values of manufacturer
  geom_histogram(binwidth=2, alpha=0.5, position="identity") # position="identity" -> overlap


# Interlayer histograms
ggplot(mpg_subset, aes(x=hwy, fill=manufacturer)) +
  geom_histogram(binwidth=2, position="dodge") # position="dodge" -> interlayer


# Density curves
ggplot(mpg_subset, aes(x=hwy, fill=manufacturer)) +
  geom_density()

ggplot(mpg_subset, aes(x=hwy, fill=manufacturer)) +
  geom_density(alpha=0.4)

```



More histograms and options
```{r}
# Histograms more sofisticated 
set.seed(4235)
diamonds_small <- diamonds[sample(nrow(diamonds), 1000), ] # Get a random small subset

ggplot(diamonds_small, aes(x=price)) +
  geom_histogram(binwidth=1000) # We see its histogram 

ggplot(diamonds_small, aes(x=price, ..density.., colour=cut)) +
  geom_freqpoly(binwidth=1000) # Show frequency polinomials with a colours by a factor (cut)

ggplot(diamonds_small, aes(x=price, fill=cut)) +
  geom_histogram(binwidth=1000) # Overlapped histograms with a colours by a factor (cut)

ggplot(diamonds_small, aes(x=price, fill=cut)) +
  geom_density(alpha=0.3) # Overlapped densisty curvers with a colours by a factor (cut)

```




### Bar charts

We load a tips dataset to observ its distribution

```{r}
# install.packages("reshape2")
library(reshape2)
data(tips)
help(tips)
head(tips)
```

```{r}
# Basic bar chart of total bill respect of day time
# We can observe that at dinner, the total bill is more expensive than at lunch
ggplot(data=tips, aes(x=time, y=total_bill)) +
  geom_bar(stat="identity") 

# We plot the last bar chart by smoker
ggplot(data=tips, aes(x=time, y=total_bill, fill=smoker)) +
  geom_bar(stat="identity")

```

It can be observed that sunday is when the tables pay more of mean.
```{r}
# Count of clients by day
ggplot(data=tips, aes(x=day)) +
    geom_bar(stat="count") 

# Mean count of bills by day
ggplot(data=tips, aes(x=day, y=total_bill)) +
    geom_bar(stat = "summary", fun.y = "mean") # fun.y="mean" applies the function "mean" on Y axe
```







### Factor charts with error bars

This charts are same bar charts, but with lines which connects points. We load a dataset of tooth growth on guinea pings.

```{r}
data(ToothGrowth)
help(ToothGrowth)
head(ToothGrowth)
tg <- ToothGrowth
```

We do a variables summary which allows us to realize displays with confidence margins:

```{r}
# install.packages("Rmisc")
library(Rmisc)
tgc <- summarySE(tg, measurevar="len", groupvars=c("supp","dose")) 
# measurevar="len" -> variable of measure: "len", groupvars=c("supp","dose") -> variables to group: "supp", "dose".
tgc
```

We do factor charts. We can observe the difference to apply a supplement or another supplement to pig. The vertical bars are confidence intervals of 95% of the tooth length values according to the amount of dose contributed in each supplement.

```{r}
# Factor chart with error bars
ggplot(tgc, aes(x=dose, y=len, colour=supp)) +
  geom_errorbar(aes(ymin=len-se, ymax=len+se), width=0.1) + # It adds a error bars with ymin and ymax like len-se (standard error) len+se
  geom_line() + # It adds a line 
  geom_point() # It adds a points

# We observe how would be this chart according to tips by day
ggplot(tips, aes(x=day, y=tip, colour="red")) +
  geom_line() +
  geom_point()

```




### Boxplot charts 

```{r}
# We do the boxplot chart with diamonds dataset
ggplot(diamonds, aes(cut, price)) + 
  geom_boxplot()

# Boxplot with the tips by day
# We can observe that the sunday they pay more of mean, but the saturday there are more people with higher tips
ggplot(tips, aes(day, tip)) +  
  geom_boxplot() + 
  coord_flip() # It flips the axes
```


