---
title: "Exploratory analysis climatics data"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---

We make a graphic study of the global temperature evolution from 1850. We want to know:

- If there have been a important increasing of global temperature .
- If it is produced more anomalies than before, thermally.
- To take ideas and descriptors as for the given information.

Data are global temperatures and are described like:

- Date: starts in 1750 for average land temperature and 1850 for max and min land temperatures and global ocean and land temperatures
- LandAverageTemperature: global average land temperature in celsius
- LandAverageTemperatureUncertainty: the 95% confidence interval around the average
- LandMaxTemperature: global average maximum land temperature in celsius
- LandMaxTemperatureUncertainty: the 95% confidence interval around the maximum land temperature
- LandMinTemperature: global average minimum land temperature in celsius
- LandMinTemperatureUncertainty: the 95% confidence interval around the minimum land temperature
- LandAndOceanAverageTemperature: global average land and ocean temperature in celsius
- LandAndOceanAverageTemperatureUncertainty: the 95% confidence interval around the global average land and ocean temperature



### Let's study the dataset

We load data and show its information:
```{r}
data <- read.csv("./data/global-temperatures-data.csv")
head(data)
```

Like we can observe, data contains missing values (`NA`). For this, we are going to remove all rows which contains missing values:
```{r}
data <- na.omit(data)
head(data)
```

Let's see how the columns data are:
```{r}
str(data)
```

The date of data are factor, therefore, we are going to convert this factor to date:
```{r}
data$dt <- as.Date(data$dt, "%Y-%m-%d")
str(data)
```

Now, we extract the month and year from date and we add it to data:
```{r}
data$Month <- format(data$dt, "%m")
data$Month.String <- format(data$dt, "%B")
data$Year <- format(data$dt, "%Y")
head(data)
```


### Let's represent the average temperatures by month from 1850

```{r}
# Load the necessary libraries:
library(ggplot2)
library(dplyr)

# Represent the average temperatures
ggplot(data, aes(x = dt, y = LandAverageTemperature, colour = reorder(Month.String, -LandAverageTemperature, mean))) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Land Average Temperature by month", x = "Year", y = "Temperature", colour = "Month")

ggplot(data, aes(x = dt, y = LandAndOceanAverageTemperature, colour = reorder(Month.String, -LandAndOceanAverageTemperature, mean))) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Land and Ocean Average Temperature by month", x = "Year", y = "Temperature", colour = "Month")
```

It is observed that while the land average temperature has increased smoothly, the land and ocean average temperature has increased more sharply.



### Let's represent the differences between the land maximum and minimum temperatures

First, we are going to the maximum and minimum temperatures evolution over time:
```{r}
# Let's represent the mean of maximum temperatures by year
max_by_year <- data %>%
                filter(Year >= 1850) %>%
                group_by(Year) %>%
                summarize(Temperature = mean(LandMaxTemperature))

ggplot(max_by_year, aes(x = Year, y = Temperature, colour = Temperature)) +
  geom_point() +
  geom_smooth(method = "loess") +
  ggtitle("Max temperatures of the land by year") +
  scale_colour_gradient(low = "blue", high = "red") +
  scale_x_discrete(breaks = scales::pretty_breaks(n = 10))
```


```{r}
# Let's represent the mean of minimum temperatures by year
min_by_year <- data %>%
                filter(Year >= 1850) %>%
                group_by(Year) %>%
                summarize(Temperature = mean(LandMinTemperature))

ggplot(min_by_year, aes(x = Year, y = Temperature, colour = Temperature)) +
  geom_point() +
  geom_smooth(method = "loess") +
  ggtitle("Min temperatures of the land by year") +
  scale_colour_gradient(low = "blue", high = "red") +
  scale_x_discrete(breaks = scales::pretty_breaks(n = 10))
```

We can observe that as the maximum as the minimum temperatures have increased 3ºC from 1850.


Now, we are going to observe the difference between maximum and minimum temperatures:
```{r}
# We add a new column with differences between maximum and minimum temperatures:
data$MaxMinLandDifferences <- data$LandMaxTemperature - data$LandMinTemperature

ggplot(data, aes(x = dt, y = MaxMinLandDifferences, colour = MaxMinLandDifferences)) + 
  geom_point() +
  geom_smooth(method = "loess") +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Differences between maximum and minimum temperatures", 
       x = "Year", 
       y = "Differences",
       colour = "Differences")
```

Also, we are going to observe this difference by quarter evolution:
```{r}
# We add a new colum with quarter of year:
data$Quarter <- quarters(data$dt)

# We represent the difference by quarter:
ggplot(data, aes(x = dt, y = MaxMinLandDifferences, color = reorder(Quarter, -MaxMinLandDifferences, mean))) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Differences between maximum and minimum temperatures by quarter",
       x = "Year",
       y = "Difference",
       colour = "Quarter")
```

Like we can observe, over time it is not seem a difference in evolution of the differences between the land maximum and minimum temperatures and either throught the quarters.



### Let's see the evolution of uncertainty above the land and the ocean temperatures

```{r}
head(data)
ggplot(data, aes(x = dt, y = LandAndOceanAverageTemperatureUncertainty, colour = LandAndOceanAverageTemperatureUncertainty)) +
  geom_point() +
  geom_smooth(method = "loess") +
  scale_color_gradient() +
  labs(title = "Average of uncertainty above the land and the ocean temperatures over time",
       x = "Year",
       color = "")
```

With this, we can observe how the uncertainty of temperatures has been decreasing over time. This means that the measurements have improve a lot.



### Let's see the land and the ocean temperature distribution throught the year

We are going to represent data with boxplot chart:
```{r}
ggplot(data, aes(x = Month, y = LandAndOceanAverageTemperature)) +
  geom_boxplot(fill = "gray")
```

With this, we can observe that almost all months have the same anomalies or outliers, both above and below.


Now, we are going to represent the temperatures distribution by month throught the selected years:

```{r}
selected_years <- data[data$Year %in% c(1850, 1900, 1950, 2000, 2005), ]

ggplot(selected_years, aes(x = Month, y = LandAndOceanAverageTemperature, colour = reorder(Year, -LandAndOceanAverageTemperature, mean))) +
  geom_point() +
  scale_color_discrete("") +
  labs(title = "Average temperature in the land and the ocean by month of selected years",
       x = "Month",
       y = "Average temperature")
```
