---
title: "Probability and Statistical Inference"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---



### Section A

We have the IQ of a sample of 5 subjects:
 110
 100
 115
 105
 104

We want to:
  
##### **- Case 1: Calculate the confidence interval above the mean (95%).**
  
```{r}
iq <- c(110,100,115,105,104)
# Calculate the confidence interval
t.test(iq)
```
  
  The confidence interval (95%) is: [99.59193, 114.00807].

##### **- Case 2: Indicate if this subjects have been extracted of a population with mean = 100.**
    
```{r}
# Calculate if the mean is equal to 100
t.test(iq, mu=100, conf.level = 0.99)
```

  As the `p-value > 0.05`, then it is supossed that the subjects's population has mean = 100.



### Section B

With the dataset `survey`, we want to contrast if the pulse difference between men and women is different or not. Could it considered that the women pulse is higher than the men pulse with a confidence level of 90%?

```{r}
library(MASS)
head(survey)
```

The dataset contains `NA` values. We omit them:
```{r}
data <- na.omit(survey)

# Let's see the difference in means:
t.test(Pulse ~ Sex, data, conf.level = 0.9)
t.test(Pulse ~ Sex, data, alternative = "g", conf.level = 0.90)
```

With these results, the difference between the pulse mean of both sex is not 0 and either greater 0. Therefore, we cannot say if the women pulse is greater than the men pulse.



### Section C

**Ley de Benford**
The Xdata company has sent me a bills which seem to be false. These bills has a lot of numbers which does not seem to generated with a natural mode. We check if this bill is generated artificially with the Benford law (https://es.wikipedia.org/wiki/Ley_de_Benford). The Benford law sets that the natural distribution of the first digits is:

      0.301,0.176,0.125,0.097,0.079,0.067,0.058,0.051,0.046

The frequencies of the first digits of the company bills are:

      7, 13, 12,  9,  9, 13, 11, 10, 16

¿Are the bills falses?

```{r}
# Probabilities of the first digits (Benford)
prob <- c(0.301,0.176,0.125,0.097,0.079,0.067,0.058,0.051,0.046)
# Frequencies of the first digits (Xdata bills)
freq <- c(7, 13, 12,  9,  9, 13, 11, 10, 16)

# Check if they are dependients with probabilities
chisq.test(freq, p = prob)
```

As the `p-value < 0.05`, the distributions cannot be considered the same. Thus, the bills do not comply with Benford law and are potentially falses.





### Section D

With this dataset:
```{r}
data("PlantGrowth")
PlantGrowth
```

##### **- Could it considered that with the three treatments, the plants have the same growth?**

```{r}
# We are going to a exploratory analysis for seeing its behaviour:
library(ggplot2)
ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot()

ggplot(PlantGrowth, aes(x = weight, fill = group)) +
  geom_histogram()
```
  
  As we can see, the three treatments have not the same growth above the plants.
  
##### **- Make the one-way ANOVA. What could conclusions infer of this way?**
  
  Let's see if the groups are not equals
```{r}
oneway.test(weight ~ group, PlantGrowth)
```

  As the groups are not equals, let's see if there is any difference between `ctrl` and `trt1` treatments
```{r}
ctrl_trt1 <- PlantGrowth[PlantGrowth$group != "trt2", ]
ctrl_trt1
oneway.test(weight ~ group, ctrl_trt1)
```

With this result, we can say that the `ctrl` and `trt1` are the same. We are going to make it sure.

```{r}
new_data <- PlantGrowth
new_data[new_data$group == "ctrl", "group"] <- "trt1"
new_data
oneway.test(weight ~ group, new_data)
```

With this: `p-value < 0.05`, we make sure that there are only two treatments: `trt1` and `trt2`.