---
title: "Consumer Behaviour Analytics Tutorial - Guided Exercise"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---


```{r}
# Clear all object
rm(list=ls())
# Set working directory
setwd("/home/jmssalas/git/master-bigdata-businessanalytics/08-analytical-applications/03-location-analytics")
```


```{r}
# Import needed libraries
is.installed <- function(package) is.element(package, installed.packages())

if (!is.installed('rgdal')) 
  install.packages('rgdal', dependencies = T)

if (!is.installed('rgeos')) 
  install.packages('rgeos', dependencies = T)

library(rgdal)
library(rgeos)
```




## Read Madrid Data
```{r}
i_data_hotels1 <- readOGR(dsn = "data", layer = "hoteles1", encoding = "latin1")
i_data_hotels2 <- readOGR(dsn = "data", layer = "hoteles2", encoding = "latin1")
i_data_hotels3 <- readOGR(dsn = "data", layer = "hoteles3", encoding = "latin1")
i_data_hotels4 <- readOGR(dsn = "data", layer = "hoteles4", encoding = "latin1")
i_data_hotels5 <- readOGR(dsn = "data", layer = "hoteles5", encoding = "latin1")
i_data_turismoffice <- readOGR(dsn = "data", layer = "oficinaturismo", encoding = "latin1")
i_data_districts <- readOGR(dsn = "data", layer = "200001543", encoding = "latin1")
```

---

## Data Preprocessing

#### Union of hotels spatial data
```{r}
i_data_hotels <- rbind(i_data_hotels1,
                       i_data_hotels2,
                       i_data_hotels3,
                       i_data_hotels4,
                       i_data_hotels5)
i_data_hotels
```

---

#### Project data using EPSG code of Madrid
```{r}
EPSG <- make_EPSG()
EPSG[grepl("Spain$", EPSG$note),]
```


```{r}
# Project using the code 2062
proj4string(i_data_hotels) <- CRS("+init=epsg:2062")
proj4string(i_data_turismoffice) <- CRS("+init=epsg:2062")
proj4string(i_data_districts) <- CRS("+init=epsg:2062")
```


```{r}
i_data_hotels <- spTransform(i_data_hotels, proj4string(i_data_hotels))
i_data_hotels@data
```


```{r}
i_data_turismoffice <- spTransform(i_data_turismoffice, proj4string(i_data_turismoffice))
i_data_turismoffice@data
```


```{r}
i_data_districts <- spTransform(i_data_districts, proj4string(i_data_districts))
i_data_districts@data
```

---

#### Select districts of center of Madrid
```{r}
i_data_center <- i_data_districts[as.character(i_data_districts$GEOCODIGO) >= "079011" & as.character(i_data_districts$GEOCODIGO) <= "079052", ]

# Initial visualization
plot(i_data_center)
points(i_data_hotels, cex = 0.5)
points(i_data_turismoffice, cex = 1, pch = 19)
```


---

#### Spatial join to get only the turism offices that stay in the center districts of Madrid
```{r}
i_data_hotels_center <- i_data_hotels[i_data_center,]
i_data_turismoffice_center <- i_data_turismoffice[i_data_center,]

# Visualization
plot(i_data_center)
points(i_data_hotels_center, cex = 0.5)
points(i_data_turismoffice_center, cex = 1, pch = 19)
```



---

## Insight Collecting

#### Influence zone of current turism offices of center of Madrid collection
```{r}
# Influence zone of turism office: 500 metres
i_inf_turismoffice <- gBuffer(spgeom = i_data_turismoffice_center, width = 500)

# Visualization
plot(i_data_center)
plot(i_inf_turismoffice, add = T)
points(i_data_hotels_center, cex = 0.5)
points(i_data_turismoffice_center, cex = 1, pch = 19)
```

---

#### Spatial Filtering
```{r}
# Let's get only the hotels into influence zone
i_data_hotels_center_office <- i_data_hotels[i_inf_turismoffice, ]

# Visualization
plot(i_data_center)
plot(i_inf_turismoffice, add = T)
points(i_data_hotels_center, cex = 0.5)
points(i_data_hotels_center_office, cex = 0.1)
# points(i_data_turismoffice_center, cex = 1, pch = 19)
```


```{r}
# Let's get only the hotels out the influence zone
sel_label_turismoffice <- i_data_hotels_center_office$ETIQUETA
sel_label_no_turismoffice <- is.element(i_data_hotels_center$ETIQUETA, sel_label_turismoffice)
i_data_hotels_center_no_office <- i_data_hotels_center[!sel_label_no_turismoffice, ]

# Visualization
plot(i_data_center)
plot(i_inf_turismoffice, add = T)
points(i_data_hotels_center_no_office, cex = 0.5)
# points(i_data_turismoffice_center, cex = 1, pch = 19)
```


---

#### Let's make a segmentation with the coordinates of the hotels out of influence zone
```{r}
# Clustering above spatial coordinates
set.seed(44)
segmentation <- kmeans(i_data_hotels_center_no_office@coords, 4)
str(segmentation)
```

```{r}
# Centers of segmentation
segmentation$centers
```


```{r}
# Convert the centers of segmentation to spatial points and project them
segment_esp <- SpatialPoints(coords = segmentation$centers)
proj4string(segment_esp) <- CRS("+init=epsg:2062")
segment_esp <- spTransform(segment_esp, proj4string(segment_esp))
str(segment_esp)
```



---

## Results presentation
```{r}
plot(i_data_center)
plot(i_inf_turismoffice, add = T)
points(i_data_hotels_center_no_office, cex = 1, pch = 19, col = segmentation$cluster)
points(i_data_turismoffice_center, cex = 1, pch = 19)
points(segment_esp, cex = 2, pch = 19)
```